---
title: "Assignment - Data Wrangling - ESM 262"
author: "Celine Mol"
date: "5/1/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Step 1. Import Tidy 

### 1.1 Read the gazetteer data as-is (all columns; no type conversion) into a gaz_raw tibble

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(stringr)

gaz_raw <- read_delim("CA_Features_20180401.txt", delim = "|", col_names = TRUE)
```

### 1.2 Copy only the following columns into a gaz tibble (you can rename them if you like):
- feature ID
- name
- feature class
- state alpha
- county name
- primary latitude (decimal)
- primary longitude (decimal)
- source latitude (decimal)
- source longitude (decimal)
- elevation in meters
- map name
- date created
- date edited


```{r}
gaz <- gaz_raw %>% 
  select(FEATURE_ID, FEATURE_NAME, FEATURE_CLASS, STATE_ALPHA, COUNTY_NAME, PRIM_LAT_DEC, 
         PRIM_LONG_DEC, SOURCE_LAT_DEC, SOURCE_LONG_DEC, ELEV_IN_M, MAP_NAME, DATE_CREATED,
         DATE_EDITED)
```

### 1.3 Convert the gaz columns to the appropriate type. Convert any placeholders for unknown data to NA

```{r}
spec(gaz)

gaz <- type_convert(gaz, col_types = 
               cols (
                 DATE_CREATED = col_date(format = "%m/%d/%Y"),
                 DATE_EDITED = col_date(format = "%m/%d/%Y")
               ))
```

### 1.4 Delete from gaz rows where:
- the primary latitude or longitude are unknown
- the feature is not in California

```{r}
sum(gaz$PRIM_LAT_DEC == 0) 
sum(gaz$PRIM_LONG_DEC == 0)

gaz <- gaz %>%
  filter(PRIM_LAT_DEC != 0, PRIM_LONG_DEC != 0) %>% 
  filter(STATE_ALPHA == "CA")
```

### 1.5 Write the gaz tibble to a CSV file (using "|" as a delimiter).

```{r, warning=FALSE}
write_delim(gaz, "gaz", delim = "|", na = "NA")
```

## Step 2. Analyze

### 2.1 What is the most-frequently-occuring feature name?

```{r}
gaz$FEATURE_NAME <- as.factor(gaz$FEATURE_NAME)

frequent_name <- gaz %>%
  count(FEATURE_NAME) %>%
  filter(n == max(n))

paste0("The most frequently occuring feature name is ", frequent_name$FEATURE_NAME)

```


### 2.2 What is the least-frequently-occuring feature class?

```{r}
frequent_class <- gaz %>% 
  count(FEATURE_CLASS) %>% 
  filter(n == min(n))

paste0("The least frequently occuring feature classes are ", frequent_class$FEATURE_CLASS[1],
       " and ", frequent_class$FEATURE_CLASS[2])

```


### 2.3 What is the approximate center point of each county?
*Hint: Calculate the center of the bounding box of the county’s point features.*

```{r}
## For each county:
## Take the average of coordinates at lat & long
## Make a table that aggregates by county and only shows average lat long.

centerpoints <- gaz %>% 
  group_by(COUNTY_NAME) %>% 
  summarise(
    LAT = mean(PRIM_LAT_DEC),
    LONG = mean(PRIM_LONG_DEC)
  )

#kable(centerpoints, format = "latex")

centerpoints
```

### 2.4 What are the fractions of the total number of features in each county that are natural? man-made?
*Hint: Construct a tibble with two columns, one containing all possible feature classes (see “Feature Class Definitions”), and another containing the string “natural” or “man-made”, which you assign (it’s pretty obvious.) Then join this tibble to the gazetteer tibble.*

```{r, message=FALSE, warning=FALSE}
classes <- read_csv("class_definitions.csv")
colnames(classes) <- c("FEATURE_CLASS", "CLASS_DESCRIPTION")

gazetteer <- left_join(gaz, classes, by = "FEATURE_CLASS")

fractions <- gazetteer %>% 
  group_by(COUNTY_NAME) %>% 
  summarise(
    natural = sum(CLASS_DESCRIPTION == "natural"),
    manmade = sum(CLASS_DESCRIPTION == "manmade")) %>% 
  mutate(
    FRACTION_NATURAL = natural / (natural + manmade),
    FRACTION_MANMADE = manmade / (natural + manmade)) %>% 
  select(COUNTY_NAME, FRACTION_NATURAL, FRACTION_MANMADE)
 
fractions 
```



  



