---
title: "ESM262_Assignment2"
author: "Celine Mol"
date: "5/11/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1. Import Tidy 

### 1.1 Read the gazetteer data as-is (all columns; no type conversion) into a gaz_raw tibble

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(stringr)
library(DBI)

gaz_raw <- read_delim("CA_Features_20180401.txt", delim = "|", col_names = TRUE)
```

### 1.2 Copy only the following columns into a gaz tibble (you can rename them if you like):
- feature ID
- name
- feature class
- state alpha
- county name
- primary latitude (decimal)
- primary longitude (decimal)
- source latitude (decimal)
- source longitude (decimal)
- elevation in meters
- map name
- date created
- date edited


```{r}
gaz <- gaz_raw %>% 
  select(FEATURE_ID, FEATURE_NAME, FEATURE_CLASS, STATE_ALPHA, COUNTY_NAME, PRIM_LAT_DEC, 
         PRIM_LONG_DEC, SOURCE_LAT_DEC, SOURCE_LONG_DEC, ELEV_IN_M, MAP_NAME, DATE_CREATED,
         DATE_EDITED)
```

### 1.3 Convert the gaz columns to the appropriate type. Convert any placeholders for unknown data to NA

```{r}
spec(gaz)

gaz <- type_convert(gaz, col_types = 
               cols (
                 DATE_CREATED = col_date(format = "%m/%d/%Y"),
                 DATE_EDITED = col_date(format = "%m/%d/%Y")),
               na = c("Unknown", 0))
```

### 1.4 Delete from gaz rows where:
- the primary latitude or longitude are unknown
- the feature is not in California

```{r}
sum(gaz$PRIM_LAT_DEC == 0) 
sum(gaz$PRIM_LONG_DEC == 0)

gaz <- gaz %>%
  filter(PRIM_LAT_DEC != 0, PRIM_LONG_DEC != 0) %>% 
  filter(STATE_ALPHA == "CA")
```

## Step 2. Create a Database

### 2.1 Create a connection to a new gaz.db SQLite database.

```{r, warning=FALSE}

con <- DBI::dbConnect(RSQLite::SQLite(), path = ":memory:")

```

### 2.2 Copy the gaz tibble into the gaz.db database.

```{r, warning=FALSE}
copy_to(con, gaz, "gaz", overwrite = TRUE,
        temporary = FALSE,
         indexes = list(
          "FEATURE_NAME", "FEATURE_CLASS", "COUNTY_NAME",
          "PRIM_LAT_DEC", "PRIM_LONG_DEC"
        ))
```


## Step 3. Analyze

### Using SQL queries and the dbGetQuery function, create R code snippets that answer the following questions about California:

### 3.1 What is the most-frequently-occuring feature name?

```{r}

dbGetQuery(con, "SELECT MAX(COUNTS) AS MAX, FEATURE_NAME
           FROM (SELECT COUNT(FEATURE_NAME) AS COUNTS, FEATURE_NAME
           FROM gaz 
           GROUP BY FEATURE_NAME
           ORDER BY COUNT(FEATURE_NAME) DESC);")

```


### 3.2 What is the least-frequently-occuring feature class?

```{r}

dbGetQuery(con, "SELECT MIN(COUNTS) AS MIN, FEATURE_CLASS
           FROM (SELECT COUNT(FEATURE_CLASS) AS COUNTS, FEATURE_CLASS 
           FROM gaz 
           GROUP BY FEATURE_CLASS
           ORDER BY COUNT(FEATURE_CLASS));")

```


### 3.3 What is the approximate center point of each county?
*Hint: Calculate the center of the bounding box of the county’s point features.*

```{r}
## For each county:
## Tax max + min / 2 at each lat long
## Make a table that aggregates by county and only shows average lat long.

dbGetQuery(con, "SELECT COUNTY_NAME, 
                        (MAX(PRIM_LAT_DEC) + MIN(PRIM_LAT_DEC))/2 AS CENTER_PRIM_LAT_DEC,
                        (MAX(PRIM_LONG_DEC) + MIN(PRIM_LONG_DEC))/2 AS CENTER_PRIM_LONG_DEC 
           FROM gaz 
           GROUP BY COUNTY_NAME;")


```

### 3.4 What are the fractions of the total number of features in each county that are natural? man-made?
*Hint: Copy the feature class categories tibble you created into the gaz.db database. Then you can JOIN it to the the gaz table.*

```{r, message=FALSE, warning=FALSE}
classes <- read_csv("class_definitions.csv")
colnames(classes) <- c("FEATURE_CLASS", "CLASS_DESCRIPTION")

copy_to(con, classes, "classes", overwrite = FALSE,
        temporary = FALSE,
         indexes = list(
          "FEATURE_CLASS", "CLASS_DESCRIPTION"
        ))

dbGetQuery(con, "CREATE TABLE Fractions(
           FEATURE_CLASS text, 
           CLASS_DESCRIPTION text, 
           COUNTY_NAME text);")

dbGetQuery(con, "INSERT INTO Fractions
           SELECT gaz.FEATURE_CLASS, CLASS_DESCRIPTION, COUNTY_NAME
           FROM gaz JOIN classes
           ON gaz.FEATURE_CLASS=classes.FEATURE_CLASS;")

dbGetQuery(con, "SELECT COUNTY_NAME, natural, manmade, total, 
                        CAST(natural AS REAL) / total AS FRACTION_NATURAL,
                        CAST(manmade AS REAL) / total AS FRACTION_MANMADE
                FROM (SELECT COUNTY_NAME, 
                        SUM(CLASS_DESCRIPTION = 'natural') AS natural, 
                        SUM(CLASS_DESCRIPTION = 'manmade') AS manmade,
                        COUNT(CLASS_DESCRIPTION) AS total
           FROM Fractions
           GROUP BY COUNTY_NAME);")

```


